\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-03-06 analysis/paper/paper.tex
NOT COMPLETE
to do:
fix casebase absoluterisk for glmnet. 
fix ??? tags
add lines to dot whisker plot
Need to removeRcolorBrewer palletes

For the third case study, we will look at regularized estimation for the
hazard function. This can be performed using a combination of case-base
sampling and regularized logistic regression. To demonstrate this
capability, we examine the Study to Understand Prognoses Preferences
Outcomes and Risks of Treatment (SUPPORT) dataset
\citep{knaus1995support}. The SUPPORT dataset tracks death for
individuals who are considered seriously ill within a hospital.
Imputation was conducted using the \pkg{mice} package in \proglang{R}
with default settings \citep{mice}. Before imputation, there were a
total of 1000 individuals and 35 variables. After imputation, we have
690 individuals. We lose individuals due to coliniarity. If a pair of
covariates are deemed colinear to a certain degree, they are excluded
from any imputations. Here, only edu income avtisst* race ph* glucose*
adlp* were imputed. Of these, 66.09\% experienced the event of interest.
A description of each variable can be found in Table \ref{tab:support1}
and a breakdown of each categorical variable in Table
\ref{tab:support2}. The original data was obtained from the Department
of Biostatistics at Vanderbilt University. The imputed dataset is
available as part of the \pkg{casebase} package.

@@

adlp          & ADL Patient Day 3                        &                 & double           & 634          \\
adls          & ADL Surrogate Day 3                      &                 & double           & 310          \\
sfdm2         &                                          & 5               & integer          & 159          \\

@@

\begin{table}[ht]
\centering
\begin{tabular}{cc}
\hline
\textbf{Variable} & \textbf{Levels}                      \\
\hline
sex      & female                                        \\
         & male                                          \\
dzgroup  & ARF/MOSF w/Sepsis                             \\
         & COPD                                          \\
         & CHF                                           \\
         & Cirrhosis                                     \\
         & Coma                                          \\
         & Colon Cancer                                  \\
         & Lung Cancer                                   \\
         & MOSF w/Malig                                  \\
dzclass  & ARF/MOSF                                      \\
         & COPD/CHF/Cirrhosis                            \\
         & Coma                                          \\
         & Cancer                                        \\
income   & under \$11k                                   \\
         & $11-$25k                                      \\
         & $25-$50k                                      \\
         & \textgreater{}\$50k                           \\
race     & white                                         \\
         & black                                         \\
         & asian                                         \\
         & other                                         \\
         & hispanic                                      \\
sfdm2    & no(M2 and SIP pres)                           \\
         & adl\textgreater{}=4 (\textgreater{}=5 if sur) \\
         & SIP\textgreater{}=30                          \\
         & Coma or Intub                                 \\
         & \textless{}2 mo. follow-up                   
\end{tabular}
\caption{A description of each level within each categorical variable in the dataset.}
\label{tab:support2}
\end{table}

As before, the first step is to visualize the incidence density. The
population-time plot for the SUPPORT data appears in Figure
\ref{fig:support-poptime}.

@@

\begin{CodeInput}
R> data(support)
R> support_popTime <- popTime(support,
R>                            time = "d.time", event = "death",percentile_number = 0.31)
R> plot(support_popTime)
\end{CodeInput}
\end{CodeChunk}

Before we proceed with the analysis, we first remove hospital death as a
covariate, as this is directly informative of death. All covariates will
be used to estimate the hazard function. The hazard function will be
used to estimate the cumulative incidence of each individual in the
study. Then, the average cumulative incidience curve for all
observations will be used for our comparison.

Our main objective is to compute the absolute risk of death for a given
set of covariates, using regularization when fitting our model. First,
we fit a smooth hazard to the data; for the sake of this example, we
opted for a linear term for time. As casebase makes use of the glmnet
package, we interact with the fitsmoothhazard.fit function using a
matrix interface, where y contains the time and event variables and x
contains all other variables we would like to include in the model. For
the SUPPORT dataset, all factor variables were converted into dummy
variables. Here, we used lasso penalization by setting alpha to 1.

@@

\begin{CodeInput}
R> xInteractions=x*y[,2]
R> colnames(xInteractions) <- paste("d.time*", colnames(xInteractions), sep = "")
R> xFull=cbind(x,xInteractions)
R> cb.ModelInter=casebase::fitSmoothHazard.fit(xFull,y,family="glmnet",time="d.time",event="death",
R>                                        formula_time = ~d.time,alpha=1,ratio=rat)
\end{CodeInput}
\end{CodeChunk}

Then, we take our model and use the absoluteRisk function to integrate
and retrieve our absolute risk function.

@@

\begin{CodeInput}
R> #Estimating the absolute risk curve using the newData parameter.
R> casebaseAbsoluteInter=casebase::absoluteRisk(cb.ModelInter,time = seq(1,max(completeData$d.time), 400),newdata = xFull , s="lambda.1se",method=c("numerical"))
R> 
R> cabI=data.frame(casebaseAbsoluteInter[,1],rowMeans(casebaseAbsoluteInter[,-c(1)]))
\end{CodeInput}
\end{CodeChunk}

We will compare this curve to a regularized version of cox regression,
and a kaplan-meier survival curve. The cox regression absolute risk
curve with lasso penalization required two extra steps to retrieve.
Coxnet from the glmnet package has tools to fit a regularized cox model,
but requires a survival object skeleton of the selected variables, so
that the survival package tools can be used to retrive the absolute
risk. We handle this by fitting a regularized cox model with the glmnet
package, and a cox model from the survival package with the selected
variables from the regularized fit. This second model serves as a
skeleton that permits the use of survival package functions. The
coefficients from the regularized model will replace the coefficients in
our skeleton. the resulting model will have the correct coefficients
when integrating, but will have an incorrect standard error. For the
purposes of this case study, we are only interested in the absolute risk
curve itself and not the standard error.

@@

\begin{CodeInput}
R> #Create u and xCox, which will be used when fitting Cox with glmnet.
R> u=survival::Surv(time = as.numeric(y[,2]), event = as.factor(y[,1]))
R> 
R> xCoxI=as.matrix(cbind(data.frame("(Intercept)"=rep(1,each=690)),xFull))
R> colnames(xCoxI)[1]="(Intercept)"
R> 
R> #hazard for cox using glmnet
R> coxNetI=glmnet::cv.glmnet(x=xCoxI,y=u, family="cox",alpha=1)
R> #convergence demonstrated in plot
R> #plot(coxNet)
R> #taking the coefficient estimates for later use
R> nonzero_covariate_cox <- predict(coxNetI, type = "nonzero", s = "lambda.1se")
R> nonzero_coef_cox <- coef(coxNetI, s = "lambda.1se")
R> #creating a new dataset that only contains the covariates chosen through glmnet.
R> #cleanCoxData<- as.data.frame(cbind(as.numeric(workingData$d.time),as.factor(workingData$death), xCox[,nonzero_covariate_cox$X1]))
R> cleanCoxDataI<-as.data.frame(cbind(as.numeric(y[,2]),as.numeric(y[,1]),xCoxI[,nonzero_covariate_cox$X1]))
R> #newDataCox<-xCox[sam,nonzero_covariate_cox$X1]
R> #fitting a cox model using regular estimation, however we will not keep it.
R> #this is used more as an object place holder.
R> coxNetFitI <- survival::coxph(Surv(time=V1,event=V2) ~ ., data = cleanCoxDataI)
R> #The coefficients of this object will be replaced with the estimates from coxNet.
R> #Doing so makes it so that everything is invalid aside from the coefficients.
R> #In this case, all we need to estimate the absolute risk is the coefficients.
R> #Std. error would be incorrect here, if we were to draw error bars.
R> coxNetFitI$coefficients<-nonzero_coef_cox@x
R> #Fitting absolute risk curve for cox+glmnet. -1 to remove intercept
R> #posCox=nonzero_covariate_cox$X1
R> #newDataCoxI=cleanCoxDataI[,posCox]
R> 
R> abcoxNetI<-survival::survfit(coxNetFitI,type="breslow",newdata=cleanCoxDataI)
R> 
R> abCNI<-abcoxNetI
R> abCNI$surv=rowMeans(abCNI$surv)
\end{CodeInput}
\end{CodeChunk}

\begin{CodeInput}
R> fullDataCox<-as.data.frame(cbind(as.numeric(y[,2]),as.numeric(y[,1]),xCoxI[,-c(1)]))
R> coxFitI <- survival::coxph(Surv(time=V1,event=V2) ~ ., data = fullDataCox)
R> abcoxFitI<-survival::survfit(coxFitI,newdata=fullDataCox)
R> abCFI<-abcoxFitI
R> abCFI$surv<-rowMeans(abCFI$surv)
R> #cb.ModelNorm=casebase::fitSmoothHazard(V2~.,time="V1",event="V2",
R>        #                                ratio=rat,data=fullDataCox)
R> #summary(cb.ModelNorm)#need to figure out how to deal with NAs
R> #casebaseAbsoluteNorm=casebase::absoluteRisk(cb.ModelNorm,time = seq(1,max(completeData$d.time), 200),newdata = newData , s="lambda.1se",method=c("numerical"))
R> #cabn=data.frame(casebaseAbsoluteNorm[,1],rowMeans(casebaseAbsoluteNorm[,-c(1)]))
R> 
\end{CodeInput}
\end{CodeChunk}

We used the survival package to calculate the kaplan-meier absolute risk
function.

\begin{CodeInput}
R> 
R> coxFitup <- survival::coxph(Surv(time=V1,event=V2) ~ ., data = Cox)
R> abcoxFitup<-survival::survfit(coxFitup,newdata=Cox)
R> 
R> #creating a surv object to be used to fit an unadjusted absolute risk curve.
R> # (Kaplan-Meier risk curve)
R> km <- survival::Surv(time = completeData$d.time, event = completeData$death)
R> abKm<-survival::survfit(km~1,type='kaplan-meier',conf.type='log')
\end{CodeInput}
\end{CodeChunk}

Now that our four cumulative incidence estimations have been calculated,
we compare them all in Figure \ref{fig:abSC}. Both coxnet and casebase
decrease the absolute risk by the end of the study, in comparison to
kaplan-meier.

\begin{CodeInput}
R> 
R> #get covariates excluding d.time
R> estimatescb=setDT(as.data.frame(coef(cb.Model)[-c(1,2),1]), keep.rownames = TRUE)[]
R> estimatescb$Model="casebase"
R> colnames(estimatescb)<-c("Coefs","Estimate","Model")
R> estimatescox=setDT(as.data.frame(coef(coxNet)[-c(1),1]), keep.rownames = TRUE)[]
R> estimatescox$Model="Cox"
R> colnames(estimatescox)<-c("Coefs","Estimate","Model")
R> estimates=rbind(estimatescb, estimatescox)
R> estimates$Estimate[estimates$Estimate==0]=NA# make it so unchosen covariates appear empty
R> library(ggstance)# for vertical dodge
R>   ggplot(estimates, aes(x = Estimate, y = Coefs, color = Model)) +
R>         geom_segment(aes(x = 0, y = Coefs, xend = Estimate, yend = Coefs),size=0.5 ) +
R>         geom_point(position=ggstance::position_dodgev(height=0.2))
R> 
\end{CodeInput}
\end{CodeChunk}

\begin{CodeInput}
R> #get covariates excluding d.time
R> library(data.table)
R> estimatescb=setDT(as.data.frame(coef(cb.ModelInter)[-c(1,2),1]), keep.rownames = TRUE)[]
R> estimatescb$Model="casebase"
R> colnames(estimatescb)<-c("Coefs","Estimate","Model")
R> estimatescox=setDT(as.data.frame(coef(coxNetI)[-c(1),1]), keep.rownames = TRUE)[]
R> estimatescox$Model="Cox"
R> colnames(estimatescox)<-c("Coefs","Estimate","Model")
R> estimatesI=rbind(estimatescb, estimatescox)
R> estimatesI$Estimate[estimatesI$Estimate==0]=NA# make it so unchosen covariates appear empty
R> library(ggstance)# for vertical dodge
R>   ggplot(estimatesI, aes(x = Coefs, y = Estimate , color = Model)) +
R>         geom_segment(aes(y = 0, x = Coefs, yend = Estimate, xend = Coefs),size=0.5, position = position_dodge(width = 0.5)) +
R>         geom_point(position = position_dodge(width = 0.5))
R>   
R>     
\end{CodeInput}
\end{CodeChunk}

@@

Our choice of modeling the log-hazard as a linear function of covariates
allows us to develop a simple computational scheme for estimation.
However, as a downside, it does not allow us to model location and scale
parameters separately like the package \pkg{flexsurv}. For example, if
we look at the Weibull distribution as parametrised in
\code{stats::pweibull}, the log-hazard function is given by
\[ \log h(t; \alpha, \beta) = \left[\log(\alpha/\beta) - (\alpha - 1)\log(\beta)\right] + (\alpha - 1)\log t,\]
where \(\alpha,\beta\) are shape and scale parameters, respectively.
Unlike \pkg{casebase}, the approach taken by \pkg{flexsurv} allows the
user to model the scale parameter as a function of covariates. Of
course, this added flexibility comes at the cost of interpretability: by
modeling the log-hazard directly, the parameter estimates from
\pkg{casebase} can be interpreted as estimates of log-hazard ratios. To
improve the flexibility of \pkg{casebase} at capturing the scale of a
parametric family, we could replace the logistic regression with its
quasi-likelihood counterpart and therefore model over- and
under-dispersion with respect to the logistic likelihood. We defer the
study of the properties and performance of such a model to a future
article.

Future work will look at some of the methodological extensions of
case-base sampling. First, to assess the quality of the model fit, we
want to study the properties of the residuals (e.g.~Cox-Snell,
martingale). More work needs to be done to understand these residuals in
the context of the partial likelihood underlying case-base sampling. The
resulting diagnostic tools would then be integrated in this package.
Also, we are interested in extending case-base sampling to account for
interval censoring. This type of censoring is very common in
longitudinal studies, and many packages (e.g. \pkg{SmoothHazard},
\pkg{survival} and \pkg{Rstpm2}) provide functions to account for it.
Again, we hope to include any resulting methodology as part of this
package.

@@

This report was generated on 2020-03-03 13:15:36 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.2 (2019-12-12)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Ubuntu 18.04.4 LTS
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/openblas/libblas.so.3
#> LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.2.20.so
#> 
#> locale:
#>  [1] LC_CTYPE=en_CA.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_CA.UTF-8        LC_COLLATE=en_CA.UTF-8    
#>  [5] LC_MONETARY=en_CA.UTF-8    LC_MESSAGES=en_CA.UTF-8   
#>  [7] LC_PAPER=en_CA.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_CA.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] pracma_2.2.9        kableExtra_1.1.0    data.table_1.12.8  
#>  [4] glmnet_3.0-1        Matrix_1.2-18       casebase_0.2.1.9001
#>  [7] survival_3.1-8      magrittr_1.5        forcats_0.4.0      
#> [10] stringr_1.4.0       dplyr_0.8.3         purrr_0.3.3        
#> [13] readr_1.3.1         tidyr_1.0.0         tibble_2.1.3       
#> [16] ggplot2_3.2.1       tidyverse_1.3.0    
#> 
#> loaded via a namespace (and not attached):
#>  [1] Rcpp_1.0.3        lubridate_1.7.4   lattice_0.20-40   assertthat_0.2.1 
#>  [5] digest_0.6.25     foreach_1.4.7     R6_2.4.1          cellranger_1.1.0 
#>  [9] backports_1.1.5   reprex_0.3.0      stats4_3.6.2      evaluate_0.14    
#> [13] httr_1.4.1        pillar_1.4.3      rlang_0.4.5       lazyeval_0.2.2   
#> [17] readxl_1.3.1      rstudioapi_0.10   rticles_0.12      rmarkdown_1.18   
#> [21] webshot_0.5.2     munsell_0.5.0     broom_0.5.2       compiler_3.6.2   
#> [25] modelr_0.1.5      xfun_0.11         pkgconfig_2.0.3   shape_1.4.4      
#> [29] mgcv_1.8-31       htmltools_0.4.0   tidyselect_0.2.5  codetools_0.2-16 
#> [33] viridisLite_0.3.0 fansi_0.4.1       crayon_1.3.4      dbplyr_1.4.2     
#> [37] withr_2.1.2       grid_3.6.2        nlme_3.1-144      jsonlite_1.6.1   
#> [41] gtable_0.3.0      lifecycle_0.1.0   DBI_1.0.0         scales_1.1.0     
#> [45] cli_2.0.2         stringi_1.4.6     fs_1.3.1          xml2_1.2.2       
#> [49] generics_0.0.2    vctrs_0.2.3       iterators_1.0.12  tools_3.6.2      
#> [53] glue_1.3.1        hms_0.5.2         yaml_2.2.1        colorspace_1.4-1 
#> [57] rvest_0.3.5       VGAM_1.1-2        knitr_1.26        haven_2.2.0
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    review /home/turgeonm/Documents/git_repos/cbpaper
#> Head:     [8804f68] 2020-02-21: Merge branch 'Jesse-Islam-master'
\end{CodeOutput}
\end{CodeChunk}

>>>>

For the third case study, we will look at regularized estimation for the
hazard function. This can be performed using a combination of case-base
sampling and regularized logistic regression. To demonstrate this
capability, we examine the Study to Understand Prognoses Preferences
Outcomes and Risks of Treatment (SUPPORT) dataset
\citep{knaus1995support}. The SUPPORT dataset tracks death for
individuals who are considered seriously ill within 5 hospitals in the
United States of America. Imputation was conducted using the \pkg{mice}
package in \proglang{R} with default settings \citep{mice}. Before
imputation, there were a total of 9105 individuals and 47 variables. Of
those variables, a few were removed before imputation. We removed three
response variables: hospital charges, patient ratio of costs to charges
and patient micro-costs. Next, we removed hospital death as it was
directly informative of our event of interest, death. We also removed
functional disability and income as they are ordinal covariates, because
ordinal variables are treated as factors in our current model. Finally,
we removed 8 covariates relating to the results of previous findings: we
removed SUPPORT day 3 physiology score (sps), APACHE III day 3
physiology score (aps), SUPPORT model 2-month survival estimate, SUPPORT
model 6-month survival estimate, Physician's 2-month survival estimate
for pt., Physician's 6-month survival estimate for pt., Patient had Do
Not Resuscitate (DNR) order and Day of DNR order (\textless{}0 if before
study). Of these, sps and aps were added on after imputation, as they
were missing only 1 observation and permitted us to compare findings
between our model and their's.

First we impute manually using recommended normal values for
physiological measures \citep{harrell_2020}. Next, we impute a single
dataset using \pkg{mice} with default settings. After imputation, we
noted that the covariate for surrogate activities of daily living was
not imputed. This is due to colinearity between the other two covariates
for activities of daily living. As such, surrogate activites of daily
living was removed from our model. After these imputations, we have a
total of 9104 observations and 30 variables, two of which are sps and
aps, which will work as a baseline for our comparison to previous
findings. This dataset is available Its worth noting that these scores
were developed for a period of 180 days, which makes our comparison over
5.56 years exploratory. Also, our single imputation introduces bias in
the data. As the goal here is purely demonstrative, this bias will be
consistent with any of the following models. As such, their comparison
is still valid, though any interpretation requires an awareness of said
bias. A description of used variables can be found in Table
\ref{tab:support1} and a breakdown of each categorical variable in Table
\ref{tab:support2}. The original data was obtained from the Department
of Biostatistics at Vanderbilt University \citep{harrell_2020}. The
imputed dataset is available as part of the \pkg{casebase} package.

@@

adlp          & Activiites of Daily Living (ADL) Patient Day 3 &           & double           & 634          \\

@@

\begin{table}[ht]
\centering
\begin{tabular}{cc}
\hline
\textbf{Variable} & \textbf{Levels}                      \\
\hline
sex      & female                                        \\
         & male                                          \\
dzgroup  & ARF/MOSF w/Sepsis                             \\
         & COPD                                          \\
         & CHF                                           \\
         & Cirrhosis                                     \\
         & Coma                                          \\
         & Colon Cancer                                  \\
         & Lung Cancer                                   \\
         & MOSF w/Malig                                  \\
dzclass  & ARF/MOSF                                      \\
         & COPD/CHF/Cirrhosis                            \\
         & Coma                                          \\
         & Cancer                                        \\
         & \textgreater{}\$50k                           \\
race     & white                                         \\
         & black                                         \\
         & asian                                         \\
         & other                                         \\
         & hispanic                                      \\
ca       & Yes                                           \\
         & No                                            \\
         & Metastatic                                    \\
\end{tabular}
\caption{A description of each level within each categorical variable in the dataset.}
\label{tab:support2}
\end{table}

As before, the first step is to visualize the incidence density. The
population-time plot for the SUPPORT data appears in Figure
\ref{fig:support-poptime}. As we can see, the incidence density is not
constant.

@@

\begin{CodeInput}
R> support<-casebase::support
R> 
R> support_popTime <- popTime(support,
R+                            time = "d.time", event = "death",percentile_number = 0.31)
R> plot(support_popTime)
\end{CodeInput}
\begin{figure}

{\centering \includegraphics{../figures/support-poptime-1}

\caption{\label{fig:support-poptime} Population-time plot for the SUPPORT data.}\label{fig:support-poptime}
\end{figure}
\end{CodeChunk}

Before generating an absolute risk curve using regularized logistic
regression, we will explore three options as a baseline comparison,
beyond a simple Kaplan-meier curve. For all the following analyses in
this case study, we will use 80\% of the observations to train our
models, while 20\% will be used to generate absolute risk curves, all of
which will be averaged to demonstrate each model's predictive ability.
We will then compare each modeling using the IPA score (???NEED
REFERENCE HERE???) We will examine three cox regression scenarios: sps
only, aps only and all covariates excluding sps and aps. The original
study compared these scores over a span of 180 days, which may not be
comparable over the full 5.56 years. Over 5.56 years,
\ref{fig:baselines} demonstrates that ???BEST MODEL BASED ON IPA??? has
the best calibration and discrimination with an IPA score of ???INSERT
ACTUAL BEST SCORE??? . ??BEST MODEL?? will be used in the following
comparison to regularized cox, regularized logistic regression and the
unadjusted Kaplan-Meier curve.

@@

\begin{CodeInput}
R> completeData<-casebase::support
R> workingCompleteData=model.matrix(death~ .-d.time,data=completeData)[,-c(1)]#remove intercept
R> #Create u and xCox, which will be used when fitting Cox with glmnet.
R> newData=workingCompleteData
R> x=workingCompleteData
R> #x and y will be used to fit the casebase model under glmnet.
R> #x=as.matrix(sparse.model.matrix(death~ .-d.time+0,data=workingData))
R> y=data.matrix(completeData[,c(2,5)])
R> 
R> fullDataCox<-as.data.frame(cbind(as.numeric(y[,2]),as.numeric(y[,1]),x))
R> train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
R> test_index <- setdiff(1:nrow(fullDataCox), train_index)
R> 
R> train<-fullDataCox[train_index,]
R> test<-fullDataCox[test_index,]
R> coxSPS<- survival::coxph(Surv(time=V1,event=V2) ~ sps, data = train)
R> abcoxSPS<-survival::survfit(coxSPS,newdata = test)
R> 
R> coxAPS<- survival::coxph(Surv(time=V1,event=V2) ~ aps, data = train)
R> abcoxAPS<-survival::survfit(coxAPS,newdata = test)
R> 
R> coxFull<- survival::coxph(Surv(time=V1,event=V2) ~ .-sps -aps, data = train)
R> abcoxFull<-survival::survfit(coxAPS,newdata = test)
R> #library(prodlim)
R> 
R> 
R> KM<- survival::coxph(Surv(time=V1,event=V2) ~ 1, data = test)
R> abKM<-survival::survfit(KM)
R> #abKM <- prodlim(Hist(V1,V2)~1,data=test) minimal difference
R> abKMcompare<-cbind(abKM$time[abKM$time<=1631],1-abKM$surv[abKM$time<=1631])
R> colnames(abKMcompare)<-c("Time","KM")
R> 
R> baseCoxComparisonData<-cbind(abcoxSPS$time[abcoxSPS$time<=1631],1-rowMeans(abcoxSPS$surv[abcoxSPS$time<=1631,]),1-rowMeans(abcoxAPS$surv[abcoxSPS$time<=1631,]),1-abKM$surv,1-rowMeans(abcoxFull$surv[abcoxSPS$time<=1631,]))
R> colnames(baseCoxComparisonData)<-c("Time","SPS","APS","Unadjusted","Full")
R> myColors <- brewer.pal(5,"Dark2")
R> FigureWidth='90%'
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{../figures/baselines-1}

\caption{\label{fig:baselines} AbsoluteRisk curves from hazard models. Pink: All covariates without SPS nor APS. Green: APS only model. Blue: SPS only model. Purple: unadjusted curve. All three models underestimate the risk in comparison to the unadjusted curve.}\label{fig:baselines}
\end{figure}
\end{CodeChunk}

To get the absolute risk following regularized cox regression, this
required that we work around limitations in existing packages. To make
use of the \pkg{survival} package, we first get our hazard coefficients
using \pkg{glmnet}'s coxnet implementation for lasso penalization
\citep{regpathcox}. Next, we fit a cox model using the survival package
with the selected variables only, from our previous fit. Finally, we
replaced the coefficients in the survival object with those from our
glmnet fit, allowing us to make use of the absolute risk tools available
in the \pkg{survival} package.

For our penalized logistic regression model, we opted for a logistic
function of time, resulting in a Weibull distribution assumption. As
casebase makes use of the glmnet package, we interact with the
fitsmoothhazard.fit function using a matrix interface, where y contains
the time and event variables and x contains all other variables we would
like to include in the model. For the SUPPORT dataset, all factor
variables were converted into dummy variables. Here, we used lasso
penalization by setting alpha to 1, leading to our smooth hazard
function, depicted in the ??? curve in \ref{fig:cs3Final}. ??? I AM
STILL PENALIZING TIME ???

We then take our hazard and use it to predict the absolute risk of each
individual in our test set, over time. These individuals are then
average resulting in the ???curve in \ref{fig:cs3Final}. ???SAME BUG IS
BACK???

@@

\begin{CodeInput}
R> ggplot(as.data.frame(ab), aes(Time)) + 
R+     geom_line(data=as.data.frame(baseCoxComparisonData),mapping=aes(y = Full, colour = "Cox"),lwd=2)+
R+   geom_line(aes(y = `Absolute risk`, colour = Model),lwd=2)+
R+   geom_step(data=as.data.frame(abKMcompare),mapping=aes(x=Time,y=KM,color="Unadjusted"),lwd=2)+
R+   scale_colour_manual(name = "Models",values = myColors[c(4,1,3,2)])+
R+   labs(y= "Probability of Death", x = "Survival time (Days)")
\end{CodeInput}
\begin{figure}

{\centering \includegraphics{../figures/unnamed-chunk-2-1}

\caption{\label{fig:baselines} HERES SOME TEXT ABT IT}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

???SHORT SPIEL ABOUT IPA BETWEEN SEMI\_PARAMETRIC AND PARAMETRIC which
leads to conclusion of best overall model. This refers to the COEFPLOT
in a following figure???

@@

\begin{CodeInput}
R> 
R> library(dotwhisker)
R> library(broom)
R> library(dplyr)
R> 
R> estimatescb=setDT(as.data.frame(coef(cbFull)[-c(1,2),1]), keep.rownames = TRUE)[]
R> #estimatescb$std.error<-0
R> estimatescb$Model="casebase"
R> colnames(estimatescb)<-c("term","estimate","model")
R> 
R> estimatescoxFull=setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
R> #estimatescoxFull$std.error<-0
R> estimatescoxFull$Model="CoxFull"
R> colnames(estimatescoxFull)<-c("term","estimate","model")
R> 
R> lolliplotDots<-rbind(estimatescb,estimatescoxNet,estimatescoxFull)
R> lolliplotDots$conf.low<-NA
R> lolliplotDots$conf.high<-NA
R> lolliplotDots$conf.high[lolliplotDots$estimate< 0]<-0
R> 
R> lolliplotDots$conf.low[lolliplotDots$estimate> 0]<-0
R> 
R> dwplot(lolliplotDots)
\end{CodeInput}

\begin{center}\includegraphics{../figures/coefplots-1} \end{center}

\end{CodeChunk}

@@

Our choice of modeling the log-hazard as a linear function of covariates
allows us to develop a simple computational scheme for estimation.
However, as a downside, it does not allow us to model location and scale
parameters separately like the package \pkg{flexsurv}. For example, if
we look at the Weibull distribution as parametrised in
\code{stats::pweibull}, the log-hazard function is given by
\[ \log \lambda(t; \alpha, \beta) = \left[\log(\alpha/\beta) - (\alpha - 1)\log(\beta)\right] + (\alpha - 1)\log t,\]
where \(\alpha,\beta\) are shape and scale parameters, respectively.
Unlike \pkg{casebase}, the approach taken by \pkg{flexsurv} allows the
user to model the scale parameter as a function of covariates. Of
course, this added flexibility comes at the cost of interpretability: by
modeling the log-hazard directly, the parameter estimates from
\pkg{casebase} can be interpreted as estimates of log-hazard ratios. To
improve the flexibility of \pkg{casebase} at capturing the scale of a
parametric family, we could replace the logistic regression with its
quasi-likelihood counterpart and therefore model over- and
under-dispersion with respect to the logistic likelihood. We defer the
study of the properties and performance of such a model to a future
article.

Future work will look at some of the methodological extensions of
case-base sampling. First, to assess the quality of the model fit, we
want to study the properties of the residuals (e.g.~Cox-Snell,
martingale). More work needs to be done to understand these residuals in
the context of the partial likelihood underlying case-base sampling. The
resulting diagnostic tools would then be integrated in this package.
Also, we are interested in extending case-base sampling to account for
interval censoring. This type of censoring is very common in
longitudinal studies, and many packages (e.g.~\pkg{SmoothHazard},
\pkg{survival} and \pkg{Rstpm2}) provide functions to account for it.
Again, we hope to include any resulting methodology as part of this
package.

@@

This report was generated on 2020-03-05 23:20:03 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.1 (2019-07-05)
#> Platform: x86_64-w64-mingw32/x64 (64-bit)
#> Running under: Windows 10 x64 (build 18363)
#> 
#> Matrix products: default
#> 
#> locale:
#> [1] LC_COLLATE=English_United States.1252 
#> [2] LC_CTYPE=English_United States.1252   
#> [3] LC_MONETARY=English_United States.1252
#> [4] LC_NUMERIC=C                          
#> [5] LC_TIME=English_United States.1252    
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.5         dotwhisker_0.5.0    RColorBrewer_1.1-2 
#>  [4] pracma_2.2.9        kableExtra_1.1.0    data.table_1.12.8  
#>  [7] glmnet_3.0-2        Matrix_1.2-17       casebase_0.2.1.9001
#> [10] survival_3.1-8      magrittr_1.5        forcats_0.5.0      
#> [13] stringr_1.4.0       dplyr_0.8.3         purrr_0.3.3        
#> [16] readr_1.3.1         tidyr_1.0.2         tibble_2.1.3       
#> [19] ggplot2_3.3.0       tidyverse_1.3.0    
#> 
#> loaded via a namespace (and not attached):
#>  [1] httr_1.4.1        VGAM_1.1-2        jsonlite_1.6.1    viridisLite_0.3.0
#>  [5] foreach_1.4.8     modelr_0.1.6      assertthat_0.2.1  stats4_3.6.1     
#>  [9] ggstance_0.3.3    cellranger_1.1.0  yaml_2.2.1        pillar_1.4.3     
#> [13] backports_1.1.5   lattice_0.20-38   glue_1.3.1        digest_0.6.25    
#> [17] rvest_0.3.5       colorspace_1.4-1  plyr_1.8.6        htmltools_0.4.0  
#> [21] pkgconfig_2.0.3   haven_2.2.0       scales_1.1.0      webshot_0.5.2    
#> [25] mgcv_1.8-28       rticles_0.14      generics_0.0.2    farver_2.0.3     
#> [29] withr_2.1.2       cli_2.0.2         crayon_1.3.4      readxl_1.3.1     
#> [33] evaluate_0.14     fs_1.3.2          fansi_0.4.1       nlme_3.1-140     
#> [37] xml2_1.2.2        tools_3.6.1       hms_0.5.3         lifecycle_0.1.0  
#> [41] munsell_0.5.0     reprex_0.3.0      compiler_3.6.1    rlang_0.4.5      
#> [45] grid_3.6.1        iterators_1.0.12  rstudioapi_0.11   labeling_0.3     
#> [49] rmarkdown_2.1     gtable_0.3.0      codetools_0.2-16  DBI_1.1.0        
#> [53] R6_2.4.1          lubridate_1.7.4   knitr_1.28        shape_1.4.4      
#> [57] stringi_1.4.6     Rcpp_1.0.3        vctrs_0.2.3       dbplyr_1.4.2     
#> [61] tidyselect_1.0.0  xfun_0.12
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master C:/Users/jesse/Documents/gits/cbpaper
#> Remote:   master @ origin (https://github.com/Jesse-Islam/cbpaper.git)
#> Head:     [843365a] 2020-03-05: In process of changing my results
\end{CodeOutput}
\end{CodeChunk}
